Bird beak's elaboration and innovation

### 2023-04-18 Analysis/07-Orthogonality_and_statistics.Rmd
Update 07-Orthogonality_and_statistics.Rmd

>>>>

## PGLS

```{r}
## Get the median estimates for elaboration and innovation
load("../Data/Processed/ploting_data.rda")
load("../Data/Processed/ploting_data_order.rda")
load("../Data/Processed/groups_median_EI.rda")
load("../Data/Processed/shapespace.rda")

## Initialise the data
model_data <- data.frame("orders" = shapespace$level2, "super_orders" = shapespace$level1)
rownames(model_data) <- rownames(shapespace) 
model_data <- cbind(model_data,
                    centroid            = ploting_data$distances_to_centre[rownames(shapespace)],
                    species_elaboration = ploting_data_order$median_elaborations[rownames(shapespace)],
                    species_innovation  = ploting_data_order$median_innovations[rownames(shapespace)],
                    orders_elaboration = NA, orders_innovation = NA)
## Attributing the median innov/elab value for each order
for(order in unique(model_data$orders)) {
    if(order != "") { # This is because a couple of species have no order
        group_num <- grep(order, names(groups_median_EI$orders$innovation))
        model_data[which(model_data$orders == order), "orders_elaboration"] <- groups_median_EI$orders$elaboration[group_num]
        model_data[which(model_data$orders == order), "orders_innovation"] <- groups_median_EI$orders$innovation[group_num]
    }
}

## Remove the tips left over from the phylogeny
trees_list <- read.tree("../Data/Raw/Tree_data/AllBirdsHackett1.tre")

run.models <- function(tree, data) {
    ## all predictors
    cat(".")
    all_preds <- phylolm::phylolm(centroid ~ species_elaboration + species_innovation + orders_elaboration + orders_innovation,
                         data = data, phy = tree, model = "lambda")
    ## macro predictors
    macro_preds <- phylolm::phylolm(centroid ~ species_elaboration + species_innovation,
                           data = data, phy = tree, model = "lambda")
    ## mega predictors
    mega_preds <- phylolm::phylolm(centroid ~ orders_elaboration + orders_innovation,
                          data = data, phy = tree, model = "lambda")
    return(list(all = all_preds, macro = macro_preds, mega = mega_preds))
}

## Run 100 models (9 sec per model - 15 minutes for 100 models)
pgls_models_out <- lapply(trees_list[sample(1:length(trees_list), size = 100)], run.models, data = model_data)
save(pgls_models_out, file = "../Data/Processed/pgls_models_out.rda")

## Summarise the models
summary_out <- lapply(pgls_models_out, lapply, summary)

## Extract the slopes, p-values and r2
extract.summary <- function(one_summary) {
    ## Get the fits
    fits <- unlist(lapply(one_summary, function(X) return(X$adj.r.squared)))
    names(fits) <- paste0(names(fits), ".r2")
    ## Get the coefficients and their p vals
    coeffs <- lapply(one_summary, function(X) return(X$coefficient))

    get.est <- function(one_coeffs) {
        vals <- c(one_coeffs[-1, c("Estimate", "p.value")])
        names(vals) <- c(paste0(rownames(one_coeffs[-1, ]), c(".slope")), paste0(rownames(one_coeffs[-1, ]), c(".pval")))
        return(vals)
    }
    return(c(fits, unlist(lapply(lapply(one_summary, function(X) return(X$coefficient)), get.est))))
}
all_estimates <- do.call(rbind, lapply(summary_out, extract.summary))

## Output table
summary_table <- data.frame(matrix(NA, ncol = 8, nrow = 8))
colnames(summary_table) <- c("model", "predictor", "median slope", "5% slope", "95% slope", "median p.val", "5% p.val", "95% p.val")
summary_table$model <- c(rep("all", 4), rep("macro", 2), rep("mega", 2))
summary_table$predictor <- rep(c("species elaboration", "species innovation", "orders elaboration", "orders innovation"), 2)
summary_table[, 3:5] <- t(apply(all_estimates[, c(4:7,12,13,16,17)], 2, quantile, probs = c(0.5, 0.05, 0.95)))
summary_table[, 6:8] <- t(apply(all_estimates[, -c(1:3,4:7,12,13,16,17)], 2, quantile, probs = c(0.5, 0.05, 0.95)))

```
